---
title: "Missing data"
output:
  distill::distill_article:
    self_contained: false
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = TRUE,
  error = TRUE
)
```

# Missing data


```{r, echo=FALSE, warning=FALSE, message=FALSE, echo=FALSE}
## Load libraries
library(here)
library(dplyr)
library(ggplot2)
library(Hmisc)
library(naniar) ## for missing data 
library(gt)
library(kableExtra)
library(DT)
library(gtExtras)
load(here::here("data", "a_crash2.rda"))
```


## Per variable missingness

Number and percentage of missing. 

```{r, message =FALSE, warning =FALSE , echo=FALSE, fig.cap="Number of missing observations per variable"}
a_crash2 %>%
  select(-time2death,-ddeath, -entryid, -trandomised, -earlydeath) %>%
  miss_var_summary() %>%
  mutate(pct_miss2 = pct_miss) %>%
  gt::gt() %>%
    gt::cols_label(
    variable =  label(variable) #md("**Variable**"),
    n_miss = md("n"),
    pct_miss = md("%"),
    pct_miss2 = md("% bar")
  ) %>%
  gt::fmt_number(
    columns = c(pct_miss),
    decimals = 2
  ) %>%
  tab_spanner(
    label = md("**Missing**"),
    columns = c(n_miss, pct_miss, pct_miss2)
  ) %>%
  gt_plt_bar_pct(column = pct_miss2, scaled = TRUE, 
                 fill = "cornflowerblue", background = "aliceblue") %>%
   cols_align("center", pct_miss2) %>%
   cols_width(4 ~ px(125)) %>%
  gt_theme_espn()

```

Display

```{r, message =FALSE, warning =FALSE , echo=FALSE}
a_crash2 %>%
  select(-time2death,-ddeath, -entryid, -trandomised, -earlydeath) %>%
  gg_miss_var(show_pct = TRUE) +
  theme_minimal(base_size = 12) +
  ggtitle("Number of missing observations per variable")
```



## Missingness patterns over variables

Each row is a patient and each column is a measurement. Black indicates a missing observation. 

```{r, message =FALSE, warning =FALSE , echo=FALSE}
a_crash2 %>%
  select(-time2death,-ddeath, -entryid, -trandomised, -earlydeath) %>%
  naniar::vis_miss(sort_miss = TRUE, show_perc_col = TRUE) +
  theme_minimal(base_size = 12) +
  ggtitle("Number of missing observations per variable")
```

Distribution of missing values per patients. 

```{r, message =FALSE, warning =FALSE , echo=FALSE}
a_crash2 %>%
  select(-time2death,-ddeath, -entryid, -trandomised, -earlydeath) %>%
  gg_miss_case()
```

Relationship of missing values per patient per observation. 

```{r, message =FALSE, warning =FALSE , echo=FALSE}
a_crash2 %>%
  select(-time2death,-ddeath, -entryid, -trandomised, -earlydeath) %>%
  gg_miss_upset()
```




## (In)complete cases

This section presents patients with a least one missing value. First we list out patients with at least one missing value in a filterable table. 

```{r, message =FALSE, warning =FALSE , echo=FALSE}
cc <-
  a_crash2 %>%  
  select(-time2death,-ddeath) %>% 
  filter(!complete.cases(.))

DT::datatable(cc)
```

Then we report the pattern of missing for this set of patients. 

```{r, message =FALSE, warning =FALSE , echo=FALSE}
cc %>% gg_miss_upset()
```



## Section session info

```{r, message =FALSE, warning =FALSE , echo=FALSE}
sessionInfo()
```
